---
title: "Rmarkdown高级应用"
author: "Leo Lee"
date: "`r Sys.Date()`"
output:
  html_document:    
    keep_md: true           # 保留生成的.md文件
    number_sections: yes    # 给每一个层级标上数字
    toc: yes                # 显示目录
    toc_depth: 3            # 最大层级
    toc_float:              # 目录悬浮
      collapsed: true       # 是否折叠
      smooth_scroll: true   # 悬浮的目录窗格是否滚动
    highlight: tango        # 代码高亮格式, 这个格式看起来不那么辣眼睛
    theme: readable         # html文档主题, 这个主题看起来还不错
    df_print: paged         # 美观的数据框打印方式
    code_folding: show      # 默认展示所有R代码
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  comment = "#>", cache = TRUE, warning = FALSE, error = TRUE)
```


I made this because I needed to do some really simple examples to get started with gt. Here are a few different gt tables that worked, using short examples with data existing in the *datasets* package or that I create here. 

我所尝试的例子: 

- 示例 0: A first shot at just making the damn thing
- 示例 1: Customizing title, column headers, and digits
- 示例 2: Updating cell colors and adding footnotes
- 示例 3: Updating NAs and hiding columns
- 示例 4: Scientific notation! 
- 示例 5: Formatted dates and column spanning labels

Overall feelings about gt: so awesome and user friendly I can hardly stand it. Make sure to check out the beautiful and wonderfully easy-to-follow documentation for the gt package: 

<https://github.com/rstudio/gt>

# 加载包
```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(gt)
```

# 打印数据框 {.tabset .tabset-fade .tabset-pills}

Option | Description
-- | --
max.print | The number of rows to print.
rows.print | The number of rows to display.
cols.print | The number of columns to display.
cols.min.print | The minimum number of columns to display.
pages.print | The number of pages to display under page navigation.
paged.print | When set to FALSE turns off paged tables.
rownames.print | When set to FALSE turns off row names.

`{r cols.print=3, rows.print=3}`控制打印的数据框的行和列数
```{r cols.print=10, rows.print=15, max.print=200}
mtcars %>% 
  mutate(
    col1 = 'a',
    col2 = 'b',
    col3 = 'c',
    col4 = 'd'
  )
```



# 示例 {.tabset}

## 示例 0 
###It really just works...

Make a table of the first 5 observations in the 'rock' dataset.


```{r}
rock %>% # Get 'rock' data
  head(5) %>% # First 5 lines only
  gt() # Make a table, it just works. 

```

## 示例 1
### The first try at customization.

Biological oxygen demand table (BOD)

Use the built-in dataset BOD to create a simple customized table. The BOD dataset (see ?BOD for documentation) has two columns: **Time** and **demand** (both numeric).

```{r}

BOD %>% # Get the data...
gt() %>% # use 'gt' to make an awesome table...
  tab_header( 
    title = "BOD Table Woooooo!", # ...with this title
    subtitle = "Hooray gt!") %>% # and this subtitle
  fmt_number( # A column (numeric data)
    columns = vars(Time), # What column variable? BOD$Time
    decimals = 2 # With two decimal places
    ) %>% 
  fmt_number( # Another column (also numeric data)
    columns = vars(demand), # What column variable? BOD$demand
    decimals = 1 # I want this column to have one decimal place
  ) %>% 
  cols_label(Time = "Time (hours)", demand = "Demand (mg/L)") # Update labels

``` 

## 示例 2 
###The one where I added colors and footnotes 

ToothGrowth (VitC effect on tooth growth in guinea pigs)

See ?ToothGrowth for dataset documentation. The ToothGrowth dataset has three variables: **len** (tooth length), **supp** (supplement ID), **dose** (dosage). I use group_by + summarize here to find mean tooth length grouped by supplement and dosage first, then create a table of the means. 

```{r}

# Some wrangling (grouped means by supplement + dosage):

tooth_length <- ToothGrowth %>% 
  group_by(supp, dose) %>% 
  summarize(
    mean_len = mean(len)
  ) %>% 
  as_tibble() 

# A gt table: 
tooth_length %>% # Take tooth_length
  gt() %>% # Make a gt table with it
  tab_header(
    title = "A title just like that", # Add a title
    subtitle = "(with something below it!)" # And a subtitle
  ) %>%
  fmt_passthrough( # Not sure about this but it works...
    columns = vars(supp) # First column: supp (character)
  ) %>% 
  fmt_number(
    columns = vars(mean_len), # Second column: mean_len (numeric)
    decimals = 2 # With 4 decimal places
  ) %>% 
  fmt_number(
    columns = vars(dose), # Third column: dose (numeric)
    decimals = 2 # With 2 decimal places
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(supp), # ...for supp column!
    colors = scales::col_factor( # <- bc it's a factor
      palette = c(
        "green","cyan"), # Two factor levels, two colors
      domain = c("OJ","VC")# Levels
  )
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(dose), # ...for dose column 
    colors = scales::col_numeric( # <- bc it's numeric
      palette = c(
        "yellow","orange"), # A color scheme (gradient)
      domain = c(0.5,2) # Column scale endpoints
  )
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(mean_len), # ...for mean_len column
    colors = scales::col_numeric(
      palette = c(
        "red", "purple"), # Overboard colors! 
      domain = c(7,27) # Column scale endpoints
  )
  ) %>% 
  cols_label(supp = "Supplement", dose = "Dosage (mg/d)", mean_len = "Mean Tooth Length") %>% # Make the column headers
  tab_footnote(
    footnote = "Baby footnote test", # This is the footnote text
    locations = cells_column_labels(
      columns = vars(supp) # Associated with column 'supp'
      )
    ) %>% 
    tab_footnote(
    footnote = "A second footnote", # Another line of footnote text
    locations = cells_column_labels( 
      columns = vars(dose) # Associated with column 'dose'
      )
    )

```

## 示例 3 
### Updating NAs, hiding columns

Using first 10 observations in the 'airquality' dataset. Existing variables: **Ozone**, **Solar.R**, **Wind**, **Temp**, **Month**, **Day** 

```{r}

airquality %>% # Take the 'airquality' dataset...
  head(10) %>% # Then just grab the first 10 observations...
  gt() %>% # Then make a gt table
  tab_header(
    title = "Air Quality dataset", # Add a title
    subtitle = "...also with a subtitle" # And a subtitle
  ) %>%
  fmt_missing( # Then reassign the 'NA' values to something new
    columns = vars(Ozone), # For variable 'Ozone', make "NA"s
    missing_text = "Teddy" # ...to "Teddy" instead
  ) %>% 
  fmt_missing( # Reassign NAs for another column
    columns = vars(Solar.R), # For variable 'Solar.R', make "NA"s
    missing_text = "Dog" # ...to "Dog" instead
  ) %>% 
  cols_hide( # Then hide the 'Wind' and 'Month' variables
    columns = vars(Wind, Month)
  )

```

## 示例 4 
### Scientific notation (...and more colors, can't help it...) 

I'll just make some data here (tibble 'fake'): 

```{r}

fake <- tribble( # Create a tibble...
  ~Species, ~Height, ~Mass, # These will be the column names (variables)
  "Blarg", 20500, 1980, # Row 1
  "Gorple", 17000, 775, # Row 2
  "Roglob", 11820, 20, # Row 3
  "Fwerpy", 8005, 352 # Row 4
) %>% 
  arrange(-Mass) # Arrange high-to-low by values in Mass column

```

Then make a little table of 'fake' with values in scientific notation:

```{r}

fake %>% 
  gt() %>% 
  tab_header(
    title = "Fake species data", # Add a title
    subtitle = "another useful subtitle" # And a subtitle
  ) %>%
  fmt_scientific( # Reformat to scientific notation...
    columns = vars(Height), # The values for the 'Height' variable
    decimals = 3 # Keeping 3 decimal places
  ) %>% 
  fmt_scientific( # Reformat to scientific notation...
    columns = vars(Mass), # The values for the 'Mass' variable
    decimals = 1 # Keeping 1 decimal place
  ) %>% 
  data_color( # Update cell colors...
    columns = vars(Mass), # ...for Mass column 
    colors = scales::col_numeric( # <- bc it's numeric
      palette = c(
        "yellow", "red", "purple"), # A color scheme (gradient)
      domain = c(1e1, 2e3) # Column scale endpoints
  )
  )
  

```

## 示例 5 
### Date formatting + column spanning labels

First, I'll make a little tibble with dates:

```{r}

date_trial <- tribble( # Create a tibble...
  ~Date, ~Temp, ~Rain, ~Flavor, # These will be the variables (column names)
  "1980-01-04", 75, 0.4, "Grape", # Row 1 
  "1975-05-28", 80, 0.2, "Strawberry", # Row 2
  "1993-09-02", 94, 0.1, "Chocolate" # Row 3
)

```

Then into a nice table: 

```{r}

date_trial %>% 
  gt() %>% 
  fmt_date(
    columns = vars(Date),
    date_style = 5 # Try different numbers here to reformat! So cool!
  ) %>% 
  tab_spanner(
    label = "Weather",
    columns = vars(Temp, Rain)
  )

```

# 复杂示例 {.tabset}


## 示例 1
```{r}
library(gt)
library(tidyverse)
library(glue)

# Define the start and end dates for the data range
start_date <- "2010-06-07"
end_date <- "2010-06-14"

# Create a gt table based on preprocessed
# `sp500` table data
sp500 %>%
  dplyr::filter(date >= start_date & date <= end_date) %>%
  dplyr::select(-adj_close) %>%
  dplyr::mutate(date = as.character(date)) %>%
  gt() %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue::glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = vars(date),
    date_style = 3
  ) %>%
  fmt_currency(
    columns = vars(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = vars(volume),
    scale_by = 1 / 1E9,
    pattern = "{x}B"
  )
```







